---
title: "Paper_plots"
author: "Ben"
date: "1/22/2019"
output: pdf_document
---

```{r libraries, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(rwebppl)
library(tidyverse)
```

```{r helpers}
# getModgetModelFittedModelData()
# ===============================
# Get cached data from model fitting.
#
# Parameters
# ----------
# modelFile: str
#  Model type.
# 
# Returns
# -------
# file.path
#    File path to model.
#
getModelFittedModelData <- function(typ) {
  dirPath <- '../data/model_predictions_data'
  file.path(dirPath, paste0("df_", typ, ".csv"))
}
```

# Analysis
```{r combine-all-models}
df_naive <- data2predictions("NAIVE_BASELINE")
df_deterministic_baseline <- deterministicBaselineDf(df_naive)
df_no_recursion <- data2predictions("NO_RECURSION")
# df_full <- data2predictions("FULL_MODEL")

df_all_models <- rbind(df_naive, df_deterministic_baseline, df_no_recursion)
```


All conditions plot
```{r}
df_all_models %>% 
    rename(`Object referrent`=type,
           `Experimental condition`=actualCondition) %>%
    mutate(modelType=ifelse(modelType=="DETERMINISTIC BASELINE", "Deterministic baseline", 
                            ifelse(modelType=="NAIVE_BASELINE", "Zipf language +\n attention", "RSA-BDA"))) %>%
    mutate(modelType=factor(modelType, levels=c("Deterministic baseline", "Zipf language +\n attention", "RSA-BDA"))) %>%
    ggplot(aes(x=actualCnt, y=estCnt, col=`Experimental condition`), alpha=0.4) +
      geom_point(aes(fill=`Experimental condition`, shape=`Object referrent`), alpha=0.6, size=3) +
      geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
      ylab('Model predicted \n counts of short form') +
      xlab('Human data \ncounts of short form') +
      theme_few() +
      facet_grid(~modelType)
```

Combined condition
```{r}
df_all_models %>% 
  filter(actualCondition=="combined") %>%
  ggplot(aes(x=actualCnt, y=estCnt, col=actualCondition), alpha=0.4) +
    geom_point(aes(fill=actualCondition, shape=type), alpha=0.6, size=3) +
    geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
    ylab('Baseline model \n counts of short form') +
    xlab('Human data \n counts of short form') +
    theme_few() +
    facet_grid(~modelType)
```

Correlations
```{r}
df_all_models %>% 
  group_by(modelType) %>%
  summarise(r2=cor(estCnt, actualCnt)^2)

df_all_models %>% 
  group_by(modelType, actualCondition) %>%
  summarise(r2=cor(estCnt, actualCnt)^2)
```


```{r}
# 20190122 BP lesion test
df_frequent <- df_bda %>%
  select(IP, actualCondition, actualFrequent, posteriorPredictiveFrequent, inferredAttention) %>%
  mutate(type='frequent') %>%
  rename(actualCnt=actualFrequent,
         estCnt=posteriorPredictiveFrequent)
  
df_infrequent <- df_bda %>% 
  select(IP, actualCondition, actualInfrequent, posteriorPredictiveInfrequent, inferredAttention) %>%
  mutate(type='infrequent') %>%
  rename(actualCnt=actualInfrequent,
         estCnt=posteriorPredictiveInfrequent)

df_predictives2 <- rbind(df_frequent, df_infrequent)
cor(df_predictives2$actualCnt, df_predictives2$estCnt)^2
```


## Subject-level inferred params?

#### alpha plot
```{r}
df_alphas <- rbind(df_naive, df_no_recursion)

df_alphas %>%
  ggplot(aes(inferredAttention, col=modelType, fill=modelType)) +
    geom_density(alpha=0.3) +
    theme_few()

```

#### all params
```{r}
df_RSA <- read.csv(getModelFittedModelData("NO_RECURSION"))

plot_alpha <- df_RSA %>%
  group_by(alpha) %>%
  summarise(n=n()) %>%
  ggplot(aes(x=alpha, y=n)) +
    geom_bar(stat='identity') +
    xlab("Alpha") +
    ylab("Count") +
    theme_few()

plot_attention <- df_RSA %>%
  ggplot(aes(inferredAttention)) +
    geom_density(alpha=0.3, fill='blue') +
    xlab("Attention") +
    ylab("Density") +
    theme_few()
  
plot_condition <- df_RSA %>%
  mutate(inferredCondition=as.character(inferredCondition)) %>%
  mutate(inferredCondition=ifelse(inferredCondition=='conventionalized', 'other', inferredCondition)) %>%
  group_by(inferredCondition) %>%
  summarise(count=n()) %>%
  ggplot(aes(x=inferredCondition, y=count)) +
    geom_bar(stat='identity', position='dodge') +
    xlab("Condition") +
    ylab("Count") +
    theme_few() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

bottom_row <- plot_grid(plot_condition, plot_alpha, labels = c('B', 'C'), align = 'h', rel_widths = c(1, 1.3))
plot_all_params <- plot_grid(plot_attention, bottom_row, labels = c('A', ''), ncol = 1, rel_heights = c(1, 1.2))
# ggsave('../../paper/figs/inferred-params-plot.png', plot=plot_all_params, width=6, height=4)
```

# Save below
-------------

### Proportions Analysis

```{r}
df_cors <- df_predictives2 %>%
  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
                         ifelse(condition=='combined' & type=='infrequent', 0,
                                ifelse(condition=='accuracy' & type=='frequent', 0,
                                       ifelse(condition=='accuracy' & type=='infrequent', 0,
                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
  gather('model', 'pred', c(baseline, estCnt)) %>%
  group_by(IP, condition, actualCnt, model) %>%
  summarise(cntShortForms=sum(pred)) %>%
  group_by(IP, model, condition) %>%
  summarise(estCnts=sum(cntShortForms),
            actCnts=sum(actualCnt)) %>%
  mutate(condition_n=ifelse(condition=='combined' | condition=='accuracy', 32, 64)) %>%
  mutate(estProp=estCnts/condition_n,
         actProp=actCnts/condition_n) %>%
  ungroup
```

```{r}
df_cors %>%
  group_by(model) %>%
  summarise(n=n(),
            r=cor(estProp, actProp)^2)
```

## Proportions plot1

```{r}
df_cors %>%
  mutate(model=ifelse(model=='baseline', 'Baseline', 'BDA RSA')) %>%
  ggplot(aes(x=actProp, y=estProp, col=condition), alpha=0.4) +
    geom_point(aes(fill=condition), alpha=0.6, size=3) +
    geom_abline(intercept=0, slope=1, col='grey', alpha=0.4) +
    # geom_point(aes(x=actualInrequentProps, y=estInfrequentProps, col=condition), alpha=0.4) +
    facet_wrap(~model) +
    ylab('Baseline model \n counts of short form') +
    xlab('Human data \n counts of short form') +
    ylim(0, 1) +
    xlim(0, 1) +
    theme_few()

ggsave('../../paper/figs/bda-vs-baseline-plot.png', plot = last_plot(), width=6, height=3)
```




Some kind of significance test?

```{r}
fisher.z <- function(r1,r2,n1,n2) {
  # Todo: review (https://stats.stackexchange.com/questions/64152/are-two-pearson-correlation-coefficients-different)
  ((0.5*log((1+r1)/(1-r1)))-(0.5*log((1+r2)/(1-r2))))/((1/(n1-3))+(1/(n2-3)))^0.5
}
  
  

```


```{r}
rs <- df_predictives2 %>%
  mutate(baseline=ifelse(condition=='combined' & type=='frequent', 24,
                         ifelse(condition=='combined' & type=='infrequent', 8,
                                ifelse(condition=='accuracy' & type=='frequent', 0,
                                       ifelse(condition=='accuracy' & type=='infrequent', 24,
                                              ifelse(condition=='time' & type=='frequent', 48, 16)))))) %>%
  filter(type=='frequent') %>%
  gather(model, value,  c(estCnt, baseline)) %>%
  select(actualCnt, model, value, condition) %>%
  group_by(model) %>%
  summarise(n=n(),
            r=cor(actualCnt, value)^2)
print(rs)
```





How well do we infer the strategies?
```{r}
mean(df_predictives$inferredCondition == df_predictives$actualCondition)
```

We use conventionatlized only for a few. These are very strange in that they *almost* always avoid using the short form for the frequent term, but use it for the non-frequent 
```{r}
df_predictives %>%
  filter(inferredCondition=='conventionalized')
```


Output inferred subject-level params
```{r}
df_predictives %>%
  select(IP, alpha, depth, inferredCondition) %>%
  toJSON() %>%
  write_lines('subjectInferredParams.json')
```


# Try to model time dynamics

```{r}
f_temporal_bda_model <- "../models/bdaFig5.wppl"
model_temporal_bda <- getModelFile(f_temporal_bda_model)
df_trajectories <- rwebppl::webppl(model_temporal_bda)
```



```{r}

df2 <- data.frame()
for (i in seq(1, length(df_trajectories))) {
  df2 <- rbind(df2, df_trajectories[[i]])
}

```
```{r}
numIPs <- length(unique(df2$IP))
df2$trialNum <- rep(seq(1, 31), numIPs)
```

```{r}
df2 %>%
  filter(trialNum==1 | trialNum==31) %>%
  spread(trialNum, data) %>%
  rename('first'=`1`, 'last'=`31`) %>%
  mutate(diff=last-first) %>%
  ggplot(aes(x=IP, y=diff)) +
    geom_bar(stat='identity', position='dodge') +
    theme_few()
```


```{r}
df2 %>%
  group_by(trialNum) %>%
  summarise(n=n(),
            avgProp=mean(data),
            ci_high=avgProp + 1.96 *sqrt(avgProp*(1-avgProp)/n),
            ci_low=avgProp - 1.96 *sqrt(avgProp*(1-avgProp)/n)) %>%
  ggplot(aes(x=trialNum, y=avgProp)) +
    geom_line() +
    geom_errorbar(aes(ymax=ci_high, ymin=ci_low)) +
    ylim(0, 1) +
    theme_few()
```

