---
title: "A pragmatic account of Zipf's Law of Abbreviation via Bayesian data analysis"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Benjamin N. Peloquin} \\ \texttt{bpeloqui@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And
    {\large \bf Noah D. Goodman} \\ \texttt{noah@university.edu} \\ Department of Computer Science \\ Stanford University
    \And {\large \bf Michael C. Frank} \\ \texttt{mike@university.edu} \\ Department of Psychology \\ Stanford University}

abstract: 
    Zipf (1935) argued that the relation between a word's frequency and length was 
    a universal property of human languages emerging from the pressure to minimize speaker-listener effort.
    Subsequent work has provided indirect evidence for a Zipfian analysis of this relation between frequency and length, what Zipf termed
    the "Law of Abbreviation". In an effort to directly isolate the impact of speaker and listener pressures,
    Kanwal et al. (2017) conducted a study using an artificial language learning paradigm. The authors showed that 
    only when both a speaker and listener pressure were present would the Law of Abbreviation emerge. The authors
    highlight that while their data is consistent with a Zipfian analysis, it is also consistent with pragmatic langage use account.
    We consider this latter account conducting a Bayesian data analysis using the data from Kanwal et al. (2017) representing
    participants as rational pragmatic agents using the Rational Speech Act framework. Fitting participant-level parameters
    we find could fit between model posterior-predictive values on the empirical data from Kanwal et al. (2017). We believe this
    analysis provides strong evidence for a connection between Zipfian notions of efficiency driven language change and Gricean 
    notions of rational-pragmatic langauge use.
    
keywords:
    "Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term."
    
output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
```


# Introduction

Zipf (1935) presented a view of human behavior that focused on  *effort minimization* as a guiding principle (Principle of Least Effort). Relying on empirical evidence from corpora, he argued that the unique dynamics of speaker- and listener-effort minimization give rise to particular distributional forms found in natural language. Zipf examined the relation between a word's frequency in a corpus and a variety of properties including its rank-frequency, its denotation size, and its length. In terms of this last property, what he later termed the "Law of Abbreviation," Zipf claimed that the more frequent words are the shorter they tend to be. This relationship has been confirmed cross-linguistically, and may also be present in animal communication systems (Ferrer-i-Cancho et, 2013).\par

Zipf argued that this relationship between frequency and length was an emergent property of competing speaker-listener pressures. Under this framing, communication can be considered a cooperative, joint activity between interlocutors. Importantly, however, this activity contains a fundamental asymmetry in terms of what is effortful for a speaker compare to what is effortful for a listener.  An optimal language, from a speaker perspective, what Zipf termed “Speaker’s economy,” would consist of a single, low-cost word which was fully ambiguous - referring to any possible referent in the world. In using such a language a listener would need to disambiguate every utterance by a speaker. By contrast, if a language were optimised solely in terms of listener effort, what Zipf termed “Auditor’s economy,” the language should bijectively map all forms to meanings so there was no need for a listener to disambiguate.\par

Zipf argued that these competing forces -- Speaker’s and Auditor’s economy -- give rise to the particular distributional forms found in all languages, including the particular relationship between word frequency and length. The issue is far from settled, however -- subsequent work has argued that such properties could also be explained by a process of random typing (Ferrer-i-Cancho & Mascoso del Pardo, 2012).  Moreover, most evidence supporting the Zipfian approach has been indirect and did not fully determine the causal role of competing speaker and listener pressures.\par

Kanwal et al. (2017) implemented an artificial language learning paradigm in order to isolate these pressures directly. In line with Zipf's original analysis, they hypothesized that only in the combined speaker and listener pressure condition would the "Law of Abbreviation" emerge. That is, participants in the combined condition should converge to a lexicon in which more frequently used words were shortened and less frequent words were longer. By contrast, they hypothesized that in conditions which there was only a speaker pressure or only a listener pressure such effects would not be seen. Results largely confirmed their hypotheses.\par

While this study represents an important experimental step in confirming Zipfian notions about lexicon-level efficiency, the authors highlight an important distinction. That is, a "distinction between a language-user's mental representation of the lexicon, and the form-meaning mappings they actually produce in communication. Participants using the short form for the frequent object and the long form for the infrequent object may still retain associations of both forms with both objects in their mental lexicon-however, the nature of the communication task in this experiment may have caused them to produce only the short form for one object and the long form for the other based on purely pragmatic considerations. Given that our experiment only recorded participants’ actual productions, we cannot with certainty distinguish between these two possible explanations for the observed behavior."\par

In the current project we examine the second explanation using a Bayesian Data Analysis in which we represent study participants as rational-pragmatic speakers and listeners using the Rational Speech act framework (Frank & Goodman, 2012; Goodman & Frank, 2016). Participant-level parameters are inferred using bayesian inference. We correlate posterior-predictive values with participant-level data found by Kanwal et al. (2017) finding high degrees of agreement with our model and their data ($r^2=0.98$). While we believe this provides compelling evidence that participants are behaving pragmatically we highlight a number of findings that our current analysis does not accomodate. Further we see these findings as possible evidence for an interpretation put forth by Kanwal et al. That is, the "pragmatics-driven asymmetry in usage may or may not lead to an immediate shift in lexical representations, it may be an important first step in such a change."\par

In the following work we introduce the empirical framework of Kanwal et al. (2017) focusing on the three-conditions most relevant to our current analysis. Subsequently, we introduce the Rational Speech Act framework (RSA), our linking function in our Bayesian data analysis. We provide a high-level overview of the particular modeling framework and the results of our analysis. We provide a discussion of the short-comings and implications of our approach ending with an example of how pragmatic driven asymmetries may drive lexical change.\par

# Experimental evidence of Zipf's Law of Abbreviation

Previous evidence for Zipf's Law of Abbreviation was ...\par

Kanwal et al. (2017) used an artificial language learning paradigm to isolate the impact of speaker- and listener-pressures. They test four conditions. In the *combined* condition two participants were paired in web experiment....\par

The authors hypothesized that a preference to map the short form to the more frequent object, but not the infrequent object should only occur in *combined* condition. By contrast, when there was no pressure to communicate, and only the speaker cost was present, participants should always prefer the shortened form, regardless of the referent. When there was cost different between the shortened and lengthened forms (speaker pressure) then participants should always prefer to use the unambiguous, long-forms regardless of referent. The authors ran an additional fourth condition in which there were no speaker or listener pressures. For the purposes of the current project we did no analyse data from this condition (REASON WHY HERE).

# A changing lexicon or pragmatic language use?

While the data is consistent with a Zipfian account of the Law of Abbreviation, Kanwal et al. (2017) highlight an important alternative explanation. They describe a distinction between the "mental representation" of the lexicon and produced forms. That is, Zipf's Law of Abbreviation is an account of lexical change, not language-use. If participants are retaining both words in their "mental lexicon," but *using the lexicon pragmatically* by using the shortened form only for the more frequent item, then the lexicon itself has not "changed" at all. Given the nature of their current task, the authors highlight that they cannot distinguish between the possible explanations. 

# Modeling framework

## Rational speech act theory as a model for pragmatic participant behavior

```{r plate-diagram, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
img <- png::readPNG("figs/plate-temp.png")
grid::grid.raster(img)
```

# Results
```{r plot-post-predictives, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/post-predictives-plot.png")
grid::grid.raster(img)
```

# Connecting pragmatic language use to language change

# Conclusion

Languages around the world display a consistent relationship between word length and frequency -- more frequent words tend to be shorter (CITATIONS).  Zipf (1935) called this particular finding the “Law of Abbreviation” and characterized it in terms of competing speaker listener pressures. While other work has provided indirect evidence consistent with this Zipfian analysis (CITATIONS), Kanwal et al. (2017) attempted to derive this effect while isolating speaker and listener pressures experimentally. In an artificial language learning paradigm the authors showed that subjects were significantly more likely to produce lexicons consistent the “Law of Abbreviation,” but only when both speaker and listener pressures are present. While this evidence is compelling the authors highlight an alternative hypothesis that may have generated the same effect, but could not be distinguished in the current study -- pragmatic language use instead of lexical change. To test this secondary hypothesis we conducted a Bayesian data analysis, modeling participants as rational pragmatic agents using the Rational Speech-act (RSA) framework. Inferring subject-level parameters we are able to fit Kanwal et al. (2017) experimental data with high-levels of accuracy ($r^2=0.986). While we believe this provides evidence for the pragmatic-language use interpretation we also highlight the compatibility of this interpretation with an additional insight of Kanwal et al. (2017) -- pragmatic language use may serve as an important bootstrap for future language change.

# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent

